The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-03 rockerverse.Rmd
only first mention uses \CRANpkg
preparing for submission, see #26

**\CRANpkg{stevedore}** is currently the most powerful Docker client in R \citep{cran_stevedore}.
It interfaces with the Docker daemon over the Docker HTTP API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}} via a Unix socket on Linux or MacOS, over a named pipe on Windows, or over an HTTP/TCP connection.
The package is the only one not using system calls to the `docker` CLI tool for managing images and containers.
The package thereby enables connections to remote Docker instances without direct configuration of the local Docker daemon.
Furthermore using the API gives access to information in a structured way, is system independent, and is likely more reliable than parsing command line output.
\CRANpkg{stevedore}'s own interface is automatically generated based on the OpenAPI specification of the Docker daemon, but it is still similar to the Docker CLI.
The interface is similar to R6 objects, in that an object of class \code{"stevedore\_object"} has a number of functions attached to it that can be called, and multiple specific versions of the Docker API can be supported thanks to the automatic generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

@@

**\pkg{containerit}** ([https://github.com/o2r-project/containerit/](https://github.com/o2r-project/containerit/)) attempts to take this one step further and includes these tools to automatically create a `Dockerfile` that can execute a given workflow \citep{nust_containerit_2019}.
\pkg{containerit} accepts an R object of classes \code{"sessionInfo"} or \code{"session\_info"} as input and provides helper functions to derive these from workflows, e.g., an R script or R Markdown document, by analysing the session state at the end of the workflow.
It relies on the \pkg{sysreqs} ([https://github.com/r-hub/sysreqs/](https://github.com/r-hub/sysreqs/)) package and it's mapping of package system dependencies to platform specific installation package names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \CRANpkg{stevedore} to streamline the user interaction and improve the created `Dockerfile`s, e.g., by running a container for the desired base image to extract the already available R packages.

@@

**\CRANpkg{liftr}** aims to solve the problem of persistent reproducible reporting in statistical computing based on the R Markdown format \citep{xie2018} for dynamic documents \citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for collaborative documents and large-scale platforms for processing documents.
\CRANpkg{liftr} makes the document the main and sole workflow control file and the only file that needs to be shared between collaborators for consistent environments, e.g. demonstrated in the DockFlow project (https://dockflow.org).
It introduces new fields to the R Markdown document header, allowing users to manually declare the dependencies, including versions, for rendering the document.
The package then generates a `Dockerfile` from this metadata and provides a utitility function to render the document inside a Docker container, i.e., \code{render\_docker("foo.Rmd")}.
An RStudio addin even allows compilation of documents with the single push of a button.

System dependencies are the domain of Docker, but for a full description of the computing environment one must also manage the R version and the the R packages.
R versions are available via the versioned Rocker image stack \citep{RJ-2017-065}.
[r-online](https://github.com/ColinFay/ronline) leverages these images and provides an app for helping users to detect breaking changes between different R versions and for historic exploration of R.
With a standalone NodeJS app or [r-online](https://srv.colinfay.me/r-online), the user can compare a piece of code run in two separate versions of R.
Internally, r-online opens one or two Docker instances with the given version of R based on Rocker images, executes a given piece of code, and returns the result to the user.
Regarding R package management, this can be achieved with MRAN, or with packages such as \CRANpkg{checkpoint} \citep{cran_checkpoint} and \CRANpkg{renv} \citep{renv2019}, which can naturally be applied within images and containers.
For example, \CRANpkg{renv} helps users to manage the state of the R library in a reproducible way, further providing isolation and portability.
While \CRANpkg{renv} does not cover system dependencies, the \CRANpkg{renv}-based environment can be transferred into a container either by restoring the environment based on the main configuration file `renv.lock` or by storing the \CRANpkg{renv}-cache on the host and not in the container \citep{ushey_using_2019}.
With both the system dependencies and R packages consciously managed in a Docker image, users can start using containers as the _only_ environment for their workflows, which allows them to work indepenently of physical computers\footnote{Allowing them to be digital "nomads", cf. J.~Bryan's \href{https://github.com/jennybc/docker-why}{https://github.com/jennybc/docker-why}.} and to assert a specific degree of confidence in the stability of a developed software \citep[cf. \texttt{README.Rmd} in][]{marwick_research_2017}.

@@

Second, the strong integration of **system libraries in core packages** in the [R-spatial community](https://www.r-spatial.org/) makes containers essential for stable and proactive development of common classes for geospatial data modelling and analysis. For example, GDAL \citep{gdal_2019} is a crucial library in the geospatial domain.
GDAL is a system dependency allowing R packages such as \CRANpkg{sf}, which provides the core data model for geospatial vector data, or \CRANpkg{rgdal}, to accommodate users to be able to read and write hundreds of different spatial raster and vector formats \citep{pebesma_simple_2018,cran_rgdal}.
\CRANpkg{sf} and \CRANpkg{rgdal} have hundreds of indirect reverse imports and dependencies and, therefore, the maintainers spend a lot of effort trying not to break them.
Purpose-built Docker images are used to prepare for upcoming releases of system libraries, individual bug reports, and for the lowest supported versions of system libraries\footnote{Cf. \href{https://github.com/r-spatial/sf/tree/master/inst/docker}{https://github.com/r-spatial/sf/tree/master/inst/docker}, \href{https://github.com/Nowosad/rspatial_proj6}{https://github.com/Nowosad/rspatial\_proj6}, and \href{https://github.com/r-spatial/sf/issues/1231}{https://github.com/r-spatial/sf/issues/1231}}.

@@

\CRANpkg{googleComputeEngineR} integrates with the R parallelisation package \CRANpkg{future} \citep{cran_future} to enable this with only a few lines or R code\footnote{\href{https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html}{https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html}}.

@@

\CRANpkg{drake} has been demonstrated to run inside containers for high reproducibility\footnote{See for example \href{https://github.com/joelnitta/pleurosoriopsis}{https://github.com/joelnitta/pleurosoriopsis} or \href{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}, the latter even running in a continuous integration platform (cf.~\nameref{development}.}.
Furthermore, \CRANpkg{drake} workflows have been shown to use \CRANpkg{future} package's function \code{makeClusterPSOCK()} for sending parts of the workflow to a Docker image for execution\footnote{\href{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}} \citep[see package's function documentation;~][]{future_2020}.
In the latter case, the container control code must be written by the user, and the \CRANpkg{future} package ensures that the host and worker can connect for communicating over socket connections.

@@

First, different packages use containers for the **deployment of R and \CRANpkg{Shiny} apps**.
\CRANpkg{Shiny} is a popular package for creating interactive online dashboards with R, and it enables users with very diverse backgrounds to create stable and user-friendly web applications \citep{cran_shiny}.
_ShinyProxy_ ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open source tool to deploy Shiny apps in an enterprise context, where it features single sign-on, but it can also be used in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarios with multiple apps.
ShinyProxy itself is written in Java to accommodate corporate requirements and may itself run in a container for stability and availability.
The tool is built on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executing long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.
Continuous integration and deployment (CI/CD) for Shiny applications using Shinyproxy can be achieved, e.g., via [GitLab pipelines](https://docs.gitlab.com/ee/ci/pipelines.html).
A pipeline can include building and checking R packages and Shiny apps.
After the code has passed the checks, Docker images are built and pushed to GitLab's container registry.
The pipeline finishes with triggering a webhook on the server, where the
deployment script is executed.
The script updates configurations, pulls the new Docker images, and restarts the services.

@@

Third, containers can greatly facilitate the **deployment to existing infrastructures**.
_Kubernetes_ ([https://kubernetes.io/](https://kubernetes.io/)) is a container-orchestration system for managing container-based application deployment and scaling.
A _cluster_ of containers, orchestrated as a single deployment, e.g., with Kubernetes, can mitigate limitations on request volumes or a container occupied with computationally intensive task.
A cluster features load-balancing, autoscaling of containers across numerous servers (in the cloud or on premise), and restarting failed ones.
Many organisations already use a Kubernetes cluster for other applications, or a managed cluster can be acquired from service providers.
Docker containers are used within Kubernetes clusters to hold native code, for which Kubernetes creates a framework around network connections and scaling of resources up and down.
Kubernetes can thereby host R applications, big parallel tasks, or scheduled batch jobs in a scalable way, and the deployment can even be triggered by changes to code repositories \citep[i.e., CD, see][]{edmondson_r_2018}.
The package \pkg{googleKubernetesR} ([https://github.com/RhysJackson/googleKubernetesR](https://github.com/RhysJackson/googleKubernetesR)) is a proof of concept for wrapping the Google Kubernetes Engine API, Google's hosted Kubernetes solution, in an easy-to-use R package.
The package \CRANpkg{analogsea} provides a way to programmatically create and destroy cloud VMs on the [Digital Ocean](https://www.digitalocean.com/) platform \citep{analogsea_2019}.
It also includes R wrapper functions to install Docker in such a VM, manage images, and control containers straight from R functions.
These functions are translated to Docker CLI commands and transferred transparently to the respective remove machine using SSH.
\CRANpkg{AzureContainers} is an umbrella package that provides interfaces for three commercial services of Microsoft's Azure Cloud, namely [Container Instances](https://azure.microsoft.com/en-us/services/container-instances/) for running individual containers, [Container Registry](https://azure.microsoft.com/en-us/services/container-registry/) for private image distribution, and [Kubernetes Service](https://azure.microsoft.com/en-us/services/kubernetes-service/) for orchestrated deployments.
While a package like \CRANpkg{plumber} provides the infrastructure for turning an R workflow into a web service, for production purposes it is usually necessary to take into account scalability, reliability and ease of management.
AzureContainers provides an R-based interface to these features and, thereby, simplifies complex infrastructure management to a number of R function calls, given an Azure account with sufficient credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.
[Heroku](https://www.heroku.com/) is another cloud platform as a service provider, and it supports container-based applications.
`heroku-docker-r` ([https://github.com/virtualstaticvoid/heroku-docker-r](https://github.com/virtualstaticvoid/heroku-docker-r)) is an independent project providing a template for deploying R applications based on Heroku's image stack, including multiple examples for interfacing R with other programming languages.
Yet the approach requires a manual management of the computing environment.

Independent integrations of R for different cloud providers lead to repeated efforts and code fragmentation.
To mitigate these problems and to avoid vendor lock-in motivated the [OpenFaaS project](https://www.openfaas.com/).
OpenFaas facilitates the deployment of functions and microservices to Kubernetes or Docker Swarm.
It is language agnostic and provides auto-scaling, metrics, and an API gateway.
Reduced boilerplate code is achieved via templates.
Template for R\footnote{See OpenFaaS R templates at \href{https://github.com/analythium/openfaas-rstats-templates}{https://github.com/analythium/openfaas-rstats-templates}.}
are provided based on Rocker's Debian and R-hub's [r-minimal](https://github.com/r-hub/r-minimal) Alpine images.
The templates use [multi-stage Docker builds](https://docs.docker.com/develop/develop-images/multistage-build/) to combine R base images with the OpenFaaS 'watchdog', a tiny Golang web server.
The watchdog marshals an HTTP request and invokes the actual application.
The R session uses \CRANpkg{plumber} or similar packages for the API endpoint with packages and data preloaded, thus minimizing response times.

@@

R has been historically viewed as a tool for analysis and scientific research, but not for creating software that corporations can rely on for production services.
However, thanks to advancements in R running as a web service, along with the ability to deploy R in Docker containers, modern enterprises are now capable of having real-time machine learning powered by R.
A number of packages and projects have enabled R to respond to client requests over TCP/IP and local socket servers, such as \CRANpkg{Rserve} \citep{cran_rserve}, \CRANpkg{svSocket} \citep{grosjean_sciviews_2019}, [rApache](http://www.rapache.net) and more recently \CRANpkg{plumber} ([https://www.rplumber.io/](https://www.rplumber.io/)) and \pkg{RestRserve} ([http://restrserve.org](http://restrserve.org)), which even processes incoming requests in parallel with forked processes using \CRANpkg{Rserve}.
The latter two also provide documentation for deployment with Docker or ready-to-use images with automated builds\footnote{See \href{https://www.rplumber.io/docs/hosting.html\#docker}{https://www.rplumber.io/docs/hosting.html\#docker}, \href{https://hub.docker.com/r/trestletech/plumber/}{https://hub.docker.com/r/trestletech/plumber/} and \href{https://hub.docker.com/r/rexyai/restrserve/}{https://hub.docker.com/r/rexyai/restrserve/}.}.
These software allow other (remote) processes and programming languages to interact with R and to expose R-based function in a service architecture with HTTP APIs.
APIs based on these packages can be deployed with scalability and high availability using containers.
This pattern of deploying code matches those used by software engineering services created in more established languages in the enterprise domain, such as Java or Python, and R can be used alongside those languages as a first-class member of a software engineering technical stack.

CARD.com implemented a web application for the optimisation of the acquisition flow and the real-time analysis of debit card transactions.
The software used \CRANpkg{Rserve} and rApache and was deployed in Docker containers.
The R session behind \CRANpkg{Rserve} acted as a read-only in-memory database, which was extremely fast and scalabale, for the many concurrent rApache processes responding to the live-scoring requests of various divisions of the company.
Similarly deodorised R scripts were responsible for the ETL processes and even the client-facing email, text message and push notification alerts sent in real-time based on card transactions.
The related Docker images were made available at [https://github.com/cardcorp/card-rocker](https://github.com/cardcorp/card-rocker).
The images extended `rocker/r-base` and additionally entailed an SSH client and a workaround for being able to mount SSH keys from the host, Pandoc, the Amazon Web Services (AWS) SDK, and Java, which is required by the AWS SDK.
The AWS SDK allowed for running R consumers reading from real-time data processing streams of [AWS Kinesis](https://aws.amazon.com/kinesis/) \footnote{See useR!2017 talk \href{https://static.sched.com/hosted\_files/user2017/2f/AWR Kinesis at useR 2017.pdf}{"Stream processing with R in AWS"}.}.
The applications were deployed on Amazon Elastic Container Service ([ECS](https://aws.amazon.com/ecs/)).
The main takeaways from using R in Docker were not only that pinning the R package versions via MRAN is important, but also that moving away from Debian testing to a distribution with long-term support can be necessary.
For the use case at hand, this switch allowed for more control over upstream updates and for minimising the risk of breaking the automated builds of the Docker images and production jobs.

The AI @ T-Mobile team created a set of neural network machine learning natural language processing models to help customer care agents manage text-based messages for customers \citep{t-mobile_enterprise_2018}.
For example, one model quickly identifies whether a message is from a customer \citep[see \CRANpkg{Shiny}-based \href{https://secure.message.t-mobile.com/v1/shiny/is-customer/app/}{demo} further described by ][]{nolis_small_2019}, and others tell which customers are likely to make a repeat purchase.
If a data scientist creates a machine learning model and exposes it through a \CRANpkg{plumber} API, then someone else on the marketing team could write software that sends different emails depending on that real-time prediction.
The models are convolutional neural networks that use the \CRANpkg{keras} package \citep{cran_keras} and run in a Rocker container.
The corresponding `Dockerfile`s are published [on GitHub](https://github.com/tmobile/r-tensorflow-api).
Since the models power tools for agents and customers, they need to have extremely high uptime and reliability.
The AI @ T-Mobile team found that the models performed well, and today these models power real-time services that are called over a million times a day.

>>>>

**\CRANpkg{stevedore}** is currently the most powerful Docker client in R \citep{cran_stevedore}.
It interfaces with the Docker daemon over the Docker HTTP API\footnote{\href{https://docs.docker.com/engine/api/latest/}{https://docs.docker.com/engine/api/latest/}} via a Unix socket on Linux or MacOS, over a named pipe on Windows, or over an HTTP/TCP connection.
The package is the only one not using system calls to the `docker` CLI tool for managing images and containers.
The package thereby enables connections to remote Docker instances without direct configuration of the local Docker daemon.
Furthermore using the API gives access to information in a structured way, is system independent, and is likely more reliable than parsing command line output.
\pkg{stevedore}'s own interface is automatically generated based on the OpenAPI specification of the Docker daemon, but it is still similar to the Docker CLI.
The interface is similar to R6 objects, in that an object of class \code{"stevedore\_object"} has a number of functions attached to it that can be called, and multiple specific versions of the Docker API can be supported thanks to the automatic generation\footnote{See \href{https://github.com/richfitz/stevedore/blob/master/development.md}{https://github.com/richfitz/stevedore/blob/master/development.md}.}.

@@

**\pkg{containerit}** ([https://github.com/o2r-project/containerit/](https://github.com/o2r-project/containerit/)) attempts to take this one step further and includes these tools to automatically create a `Dockerfile` that can execute a given workflow \citep{nust_containerit_2019}.
\pkg{containerit} accepts an R object of classes \code{"sessionInfo"} or \code{"session\_info"} as input and provides helper functions to derive these from workflows, e.g., an R script or R Markdown document, by analysing the session state at the end of the workflow.
It relies on the \pkg{sysreqs} ([https://github.com/r-hub/sysreqs/](https://github.com/r-hub/sysreqs/)) package and it's mapping of package system dependencies to platform specific installation package names\footnote{See \href{https://sysreqs.r-hub.io/}{https://sysreqs.r-hub.io/}.}.
\pkg{containerit} uses \pkg{stevedore} to streamline the user interaction and improve the created `Dockerfile`s, e.g., by running a container for the desired base image to extract the already available R packages.

@@

**\CRANpkg{liftr}** aims to solve the problem of persistent reproducible reporting in statistical computing based on the R Markdown format \citep{xie2018} for dynamic documents \citep[\href{https://nanx.me/liftr/}{https://nanx.me/liftr/}, ][]{liftr2019}.
The irreproducibility of authoring environments can become an issue for collaborative documents and large-scale platforms for processing documents.
\pkg{liftr} makes the document the main and sole workflow control file and the only file that needs to be shared between collaborators for consistent environments, e.g. demonstrated in the DockFlow project (https://dockflow.org).
It introduces new fields to the R Markdown document header, allowing users to manually declare the dependencies, including versions, for rendering the document.
The package then generates a `Dockerfile` from this metadata and provides a utitility function to render the document inside a Docker container, i.e., \code{render\_docker("foo.Rmd")}.
An RStudio addin even allows compilation of documents with the single push of a button.

System dependencies are the domain of Docker, but for a full description of the computing environment one must also manage the R version and the the R packages.
R versions are available via the versioned Rocker image stack \citep{RJ-2017-065}.
[r-online](https://github.com/ColinFay/ronline) leverages these images and provides an app for helping users to detect breaking changes between different R versions and for historic exploration of R.
With a standalone NodeJS app or [r-online](https://srv.colinfay.me/r-online), the user can compare a piece of code run in two separate versions of R.
Internally, r-online opens one or two Docker instances with the given version of R based on Rocker images, executes a given piece of code, and returns the result to the user.
Regarding R package management, this can be achieved with MRAN, or with packages such as \CRANpkg{checkpoint} \citep{cran_checkpoint} and \CRANpkg{renv} \citep{renv2019}, which can naturally be applied within images and containers.
For example, \pkg{renv} helps users to manage the state of the R library in a reproducible way, further providing isolation and portability.
While \pkg{renv} does not cover system dependencies, the \pkg{renv}-based environment can be transferred into a container either by restoring the environment based on the main configuration file `renv.lock` or by storing the \pkg{renv}-cache on the host and not in the container \citep{ushey_using_2019}.
With both the system dependencies and R packages consciously managed in a Docker image, users can start using containers as the _only_ environment for their workflows, which allows them to work indepenently of physical computers\footnote{Allowing them to be digital "nomads", cf. J.~Bryan's \href{https://github.com/jennybc/docker-why}{https://github.com/jennybc/docker-why}.} and to assert a specific degree of confidence in the stability of a developed software \citep[cf. \texttt{README.Rmd} in][]{marwick_research_2017}.

@@

Second, the strong integration of **system libraries in core packages** in the [R-spatial community](https://www.r-spatial.org/) makes containers essential for stable and proactive development of common classes for geospatial data modelling and analysis. For example, GDAL \citep{gdal_2019} is a crucial library in the geospatial domain.
GDAL is a system dependency allowing R packages such as \CRANpkg{sf}, which provides the core data model for geospatial vector data, or \CRANpkg{rgdal}, to accommodate users to be able to read and write hundreds of different spatial raster and vector formats \citep{pebesma_simple_2018,cran_rgdal}.
\pkg{sf} and \pkg{rgdal} have hundreds of indirect reverse imports and dependencies and, therefore, the maintainers spend a lot of effort trying not to break them.
Purpose-built Docker images are used to prepare for upcoming releases of system libraries, individual bug reports, and for the lowest supported versions of system libraries\footnote{Cf. \href{https://github.com/r-spatial/sf/tree/master/inst/docker}{https://github.com/r-spatial/sf/tree/master/inst/docker}, \href{https://github.com/Nowosad/rspatial_proj6}{https://github.com/Nowosad/rspatial\_proj6}, and \href{https://github.com/r-spatial/sf/issues/1231}{https://github.com/r-spatial/sf/issues/1231}}.

@@

\pkg{googleComputeEngineR} integrates with the R parallelisation package \CRANpkg{future} \citep{cran_future} to enable this with only a few lines or R code\footnote{\href{https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html}{https://cloudyr.github.io/googleComputeEngineR/articles/massive-parallel.html}}.

@@

\pkg{drake} has been demonstrated to run inside containers for high reproducibility\footnote{See for example \href{https://github.com/joelnitta/pleurosoriopsis}{https://github.com/joelnitta/pleurosoriopsis} or \href{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}{https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example}, the latter even running in a continuous integration platform (cf.~\nameref{development}.}.
Furthermore, \pkg{drake} workflows have been shown to use \pkg{future} package's function \code{makeClusterPSOCK()} for sending parts of the workflow to a Docker image for execution\footnote{\href{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}{https://docs.ropensci.org/drake/index.html?q=docker\#with-docker}} \citep[see package's function documentation;~][]{future_2020}.
In the latter case, the container control code must be written by the user, and the \pkg{future} package ensures that the host and worker can connect for communicating over socket connections.

@@

First, different packages use containers for the **deployment of R and \pkg{Shiny} apps**.
\pkg{Shiny} is a popular package for creating interactive online dashboards with R, and it enables users with very diverse backgrounds to create stable and user-friendly web applications \citep{cran_shiny}.
_ShinyProxy_ ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open source tool to deploy Shiny apps in an enterprise context, where it features single sign-on, but it can also be used in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarios with multiple apps.
ShinyProxy itself is written in Java to accommodate corporate requirements and may itself run in a container for stability and availability.
The tool is built on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executing long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.
Continuous integration and deployment (CI/CD) for Shiny applications using Shinyproxy can be achieved, e.g., via [GitLab pipelines](https://docs.gitlab.com/ee/ci/pipelines.html).
A pipeline can include building and checking R packages and Shiny apps.
After the code has passed the checks, Docker images are built and pushed to GitLab's container registry.
The pipeline finishes with triggering a webhook on the server, where the
deployment script is executed.
The script updates configurations, pulls the new Docker images, and restarts the services.

@@

Third, containers can greatly facilitate the **deployment to existing infrastructures**.
_Kubernetes_ ([https://kubernetes.io/](https://kubernetes.io/)) is a container-orchestration system for managing container-based application deployment and scaling.
A _cluster_ of containers, orchestrated as a single deployment, e.g., with Kubernetes, can mitigate limitations on request volumes or a container occupied with computationally intensive task.
A cluster features load-balancing, autoscaling of containers across numerous servers (in the cloud or on premise), and restarting failed ones.
Many organisations already use a Kubernetes cluster for other applications, or a managed cluster can be acquired from service providers.
Docker containers are used within Kubernetes clusters to hold native code, for which Kubernetes creates a framework around network connections and scaling of resources up and down.
Kubernetes can thereby host R applications, big parallel tasks, or scheduled batch jobs in a scalable way, and the deployment can even be triggered by changes to code repositories \citep[i.e., CD, see][]{edmondson_r_2018}.
The package \pkg{googleKubernetesR} ([https://github.com/RhysJackson/googleKubernetesR](https://github.com/RhysJackson/googleKubernetesR)) is a proof of concept for wrapping the Google Kubernetes Engine API, Google's hosted Kubernetes solution, in an easy-to-use R package.
The package \CRANpkg{analogsea} provides a way to programmatically create and destroy cloud VMs on the [Digital Ocean](https://www.digitalocean.com/) platform \citep{analogsea_2019}.
It also includes R wrapper functions to install Docker in such a VM, manage images, and control containers straight from R functions.
These functions are translated to Docker CLI commands and transferred transparently to the respective remove machine using SSH.
\pkg{AzureContainers} is an umbrella package that provides interfaces for three commercial services of Microsoft's Azure Cloud, namely [Container Instances](https://azure.microsoft.com/en-us/services/container-instances/) for running individual containers, [Container Registry](https://azure.microsoft.com/en-us/services/container-registry/) for private image distribution, and [Kubernetes Service](https://azure.microsoft.com/en-us/services/kubernetes-service/) for orchestrated deployments.
While a package like \pkg{plumber} provides the infrastructure for turning an R workflow into a web service, for production purposes it is usually necessary to take into account scalability, reliability and ease of management.
AzureContainers provides an R-based interface to these features and, thereby, simplifies complex infrastructure management to a number of R function calls, given an Azure account with sufficient credit\footnote{See \emph{"Deploying a prediction service with Plumber"} vignette for details:  \href{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01_plumber_deploy.html}{https://cran.r-project.org/web/packages/AzureContainers/vignettes/vig01\_plumber\_deploy.html}.}.
[Heroku](https://www.heroku.com/) is another cloud platform as a service provider, and it supports container-based applications.
`heroku-docker-r` ([https://github.com/virtualstaticvoid/heroku-docker-r](https://github.com/virtualstaticvoid/heroku-docker-r)) is an independent project providing a template for deploying R applications based on Heroku's image stack, including multiple examples for interfacing R with other programming languages.
Yet the approach requires a manual management of the computing environment.

Independent integrations of R for different cloud providers lead to repeated efforts and code fragmentation.
To mitigate these problems and to avoid vendor lock-in motivated the [OpenFaaS project](https://www.openfaas.com/).
OpenFaas facilitates the deployment of functions and microservices to Kubernetes or Docker Swarm.
It is language agnostic and provides auto-scaling, metrics, and an API gateway.
Reduced boilerplate code is achieved via templates.
Template for R\footnote{See OpenFaaS R templates at \href{https://github.com/analythium/openfaas-rstats-templates}{https://github.com/analythium/openfaas-rstats-templates}.}
are provided based on Rocker's Debian and R-hub's [r-minimal](https://github.com/r-hub/r-minimal) Alpine images.
The templates use [multi-stage Docker builds](https://docs.docker.com/develop/develop-images/multistage-build/) to combine R base images with the OpenFaaS 'watchdog', a tiny Golang web server.
The watchdog marshals an HTTP request and invokes the actual application.
The R session uses \pkg{plumber} or similar packages for the API endpoint with packages and data preloaded, thus minimizing response times.

@@

R has been historically viewed as a tool for analysis and scientific research, but not for creating software that corporations can rely on for production services.
However, thanks to advancements in R running as a web service, along with the ability to deploy R in Docker containers, modern enterprises are now capable of having real-time machine learning powered by R.
A number of packages and projects have enabled R to respond to client requests over TCP/IP and local socket servers, such as \CRANpkg{Rserve} \citep{cran_rserve}, \CRANpkg{svSocket} \citep{grosjean_sciviews_2019}, [rApache](http://www.rapache.net) and more recently \pkg{plumber} ([https://www.rplumber.io/](https://www.rplumber.io/)) and \pkg{RestRserve} ([http://restrserve.org](http://restrserve.org)), which even processes incoming requests in parallel with forked processes using \CRANpkg{Rserve}.
The latter two also provide documentation for deployment with Docker or ready-to-use images with automated builds\footnote{See \href{https://www.rplumber.io/docs/hosting.html\#docker}{https://www.rplumber.io/docs/hosting.html\#docker}, \href{https://hub.docker.com/r/trestletech/plumber/}{https://hub.docker.com/r/trestletech/plumber/} and \href{https://hub.docker.com/r/rexyai/restrserve/}{https://hub.docker.com/r/rexyai/restrserve/}.}.
These software allow other (remote) processes and programming languages to interact with R and to expose R-based function in a service architecture with HTTP APIs.
APIs based on these packages can be deployed with scalability and high availability using containers.
This pattern of deploying code matches those used by software engineering services created in more established languages in the enterprise domain, such as Java or Python, and R can be used alongside those languages as a first-class member of a software engineering technical stack.

CARD.com implemented a web application for the optimisation of the acquisition flow and the real-time analysis of debit card transactions.
The software used \pkg{Rserve} and rApache and was deployed in Docker containers.
The R session behind \pkg{Rserve} acted as a read-only in-memory database, which was extremely fast and scalabale, for the many concurrent rApache processes responding to the live-scoring requests of various divisions of the company.
Similarly deodorised R scripts were responsible for the ETL processes and even the client-facing email, text message and push notification alerts sent in real-time based on card transactions.
The related Docker images were made available at [https://github.com/cardcorp/card-rocker](https://github.com/cardcorp/card-rocker).
The images extended `rocker/r-base` and additionally entailed an SSH client and a workaround for being able to mount SSH keys from the host, Pandoc, the Amazon Web Services (AWS) SDK, and Java, which is required by the AWS SDK.
The AWS SDK allowed for running R consumers reading from real-time data processing streams of [AWS Kinesis](https://aws.amazon.com/kinesis/) \footnote{See useR!2017 talk \href{https://static.sched.com/hosted\_files/user2017/2f/AWR Kinesis at useR 2017.pdf}{"Stream processing with R in AWS"}.}.
The applications were deployed on Amazon Elastic Container Service ([ECS](https://aws.amazon.com/ecs/)).
The main takeaways from using R in Docker were not only that pinning the R package versions via MRAN is important, but also that moving away from Debian testing to a distribution with long-term support can be necessary.
For the use case at hand, this switch allowed for more control over upstream updates and for minimising the risk of breaking the automated builds of the Docker images and production jobs.

The AI @ T-Mobile team created a set of neural network machine learning natural language processing models to help customer care agents manage text-based messages for customers \citep{t-mobile_enterprise_2018}.
For example, one model quickly identifies whether a message is from a customer \citep[see \pkg{Shiny}-based \href{https://secure.message.t-mobile.com/v1/shiny/is-customer/app/}{demo} further described by ][]{nolis_small_2019}, and others tell which customers are likely to make a repeat purchase.
If a data scientist creates a machine learning model and exposes it through a \pkg{plumber} API, then someone else on the marketing team could write software that sends different emails depending on that real-time prediction.
The models are convolutional neural networks that use the \CRANpkg{keras} package \citep{cran_keras} and run in a Rocker container.
The corresponding `Dockerfile`s are published [on GitHub](https://github.com/tmobile/r-tensorflow-api).
Since the models power tools for agents and customers, they need to have extremely high uptime and reliability.
The AI @ T-Mobile team found that the models performed well, and today these models power real-time services that are called over a million times a day.
